---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


```{r Libraries}

library(tidyverse)
library(xml2)
library(XML)

```

This code takes the Voter Information Project data feed, isolates the zip urls, then downloads and extracts the XMLs to a new location. Some of the links were broken, so the original url had to be broken into pieces every time the loop stopped. 

```{r Poll Locations Data}

feed <- readLines("https://raw.githubusercontent.com/votinginfoproject/VIP-feed-information-project/master/datafeed.txt")

zip_url <- feed %>% str_subset("feed_url") %>% str_sub(14,-2)

#these were used to help identify the problem zips
zip_url1 <- zip_url[1:15] 
zip_url2 <- zip_url[18:75]
zip_url3 <- zip_url[90:157]
zip_url4 <- zip_url[159:241]

j <- 0 #counter keep your place with the problem urls
td <- tempdir()
tf <- tempfile(tmpdir=td, fileext=".zip")

for (z in zip_url) {
download.file(z, tf)
fname <- unzip(tf, list=TRUE)$Name[1]
unzip(tf, files=fname, exdir='D://PollXMLs', overwrite=TRUE)
j <- j + 1
}

```

Now it comes to parsing all of the XML files into something useful. Unfortunately lack of familiarity with XML means we don't yet have a working conversion. 

```{r Polls Data XMLs}

#this will be used for looping once a conversion method is in place
pxmls <- list.files(path = 'D://PollXMLs') 

#testing conversion with a single sample file
x <- read_xml('D://PollXMLs//Archive.zip5892875693353031553.xml')

xml_contents(xml_find_all(x, ".//election/*"))
xml_contents(xml_find_all(x, ".//locality/*"))
xml_contents(xml_find_all(x, ".//polling_location/*"))


```


These links need to be updated with the github file locations.
```{r Mail Voting Data}

census <- read.csv("MSIS2629_TeamProject-master/Exploratory_Analysis/census_with_election_data.csv")

census <- census %>% mutate(turnout = total_votes/X18.and.over.population) %>% data.frame()

vote <- read.csv("Early Voting.csv")

df <- left_join(census, vote, by=c('state.name'='State'))

df <- df %>% mutate(VoteByMail = case_when(df$No.Excuse.Absentee == 1 ~ "Any Voter", df$Absentee...Excuse.Required == 1 ~ "Excused Only", df$All.Mail.Voting == 1 ~ "All Voters"), AbsenteeStatus = case_when(df$Permanent.Absentee.Status == 1 ~ "Permanent", df$Permanent.Absentee.Status == 0 ~ "Per Case")) %>% data.frame()

df <- df[complete.cases(df),]

```

Some initial graphs to see the differences in voter turnout when vote by mail is an option.

```{r Mail Ballots Graphs}

ggplot(df, aes(turnout)) + geom_density() + facet_grid(VoteByMail~.)

ggplot(df, aes(turnout)) + geom_density() + facet_grid(AbsenteeStatus~.)


ggplot(df, aes(Percent.With.a.disability,turnout, color=VoteByMail)) + geom_smooth()

ggplot(df, aes(Median.household.income,turnout, color=VoteByMail)) + geom_smooth()

ggplot(df, aes(Poverty.Percentage,turnout, color=VoteByMail)) + geom_smooth()

ggplot(df, aes(Drive_alone,turnout, color=VoteByMail)) + geom_smooth()

ggplot(df, aes(Commute_time,turnout, color=VoteByMail)) + geom_smooth() #dig

ggplot(df, aes(Median.age..years.,turnout, color=VoteByMail)) + geom_smooth()
ggplot(df, aes(Median.age..years.,turnout, color=AbsenteeStatus)) + geom_smooth()

ggplot(df, aes(White,turnout, color=VoteByMail)) + geom_smooth()

ggplot(df, aes(African.American,turnout, color=VoteByMail)) + geom_smooth()

ggplot(df, aes(Percent.bachelor.s.degree.or.higher,turnout, color=VoteByMail)) + geom_smooth()

```

