---
title: "Voter Disenfranchisement: How laws stop eligible voters"
output:
  slidy_presentation: default
---


We will be looking at how citizens of voting age can be, and are being disenfranchised in the United States of America from these three points of view - the citizens and groups that are being adversely affected by disenfranchisement (Who), how location plays into it (the Where) and the temporal concerns that contribute to it (the When). 

```{r Libraries, include=FALSE}

library(tidyverse)

```

## Electoral and temporal concerns contributing towards disenfranchisement

Our main data source for this project comes from 2016 US Census.  We downloaded multiple files, each with various demographic information such as race, income, disability and employment. We merged these together along with results from the 2016 election to use as our main data file.

Information on early voting, abseentee ballots and other mail-in ballots came from the National Council of State Legislatures. These were html tables that we entered into .csv files to use in our analysis.

```{r Load Data, message=FALSE, warning=FALSE}

census <- read.csv("Exploratory_Analysis/census_with_election_data.csv") %>% 
    mutate(turnout = total_votes/X18.and.over.population)

laws <- read.csv("Voting Laws by State Revised.csv")

df <- left_join(census, laws, by=c('state.name'='State'))

df <- df[complete.cases(df),]

```

```{r New Vars message=FALSE, warning=FALSE}

df$EarlyVoting = case_when(df$EarlyVotingDays == 0 ~ "None",
                            df$EarlyVotingDays >= 18 ~ "18 to 45",
                            TRUE ~ "3 to 15")

df$bcollar <- df$Service + df$Natural.resources..construction..and.maintenance.occupations + df$Production..transportation..and.material.moving.occupations

df$wcollar <- df$Management..business..science..and.arts.occupations+ df$Sales.and.office.occupations

df$MailScore <- case_when(df$VoteByMail == "All Mail" ~ 1,
                         df$VoteByMail == "Any Absentee" ~ 2,
                         df$VoteByMail == "Excuse Only" ~ 3)

df$EarlyScore <- case_when(df$EarlyVoting == "None" ~ 3,
                          df$EarlyVoting == "3 to 15" ~ 2,
                          df$EarlyVoting == "18 to 45" ~ 1)

df$IDScore <- case_when(df$VoterID == "No ID" ~ 1,
                       df$VoterID == "Non Photo ID" ~ 2,
                       df$VoterID == "Strict Non-Photo" ~ 2,
                       df$VoterID == "Photo ID" ~ 3,
                       df$VoterID == "Strict Photo ID" ~ 3)

df$VoterScore <- df$MailScore + df$EarlyScore + df$IDScore

df$VoteCate <- case_when(df$VoterScore <= 4 ~ "Easy",
                         df$VoterScore >= 8 ~ "Hard",
                         TRUE ~ "Neutral")

df$DiverseCounty <- case_when(df$White <= .65 ~ "Yes",
                              TRUE ~ "No")

df$IncomeLevels <- case_when(df$Median.household.income < 40000 ~ "Low",
                             df$Median.household.income >= 55000 ~ "High",
                             TRUE ~ "Mid")

```


## County Distribution of Voter Turnout


```{r Graphs Overall Turnout Density}

ggplot(df) + geom_density(aes(x=turnout, color=VoteCate)) +
    labs(title="Flexible Voting Laws Increase Turnout",
         subtitle="County Distribution of voting in 2016 General Election",
         caption="Counties are grouped by a state level categorization on the ease of voting relative to number of early voting days, mail-in ballot rules, and voter ID laws.") +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

```


```{r Graphs Draft Turnout Density eval=FALSE, include=FALSE}

ggplot(df, aes(turnout)) + geom_density() + facet_grid(VoteByMail~.) 

ggplot(df, aes(turnout)) + geom_density() + facet_grid(factor(IDScore)~.) 

ggplot(df, aes(turnout)) + geom_density() + facet_grid(EarlyVoting~.) 

ggplot(df, aes(turnout)) + geom_density() + facet_grid(factor(VoterScore)~.) 

ggplot(df, aes(turnout)) + geom_density() + facet_grid(VoteCate~.) 

ggplot(df) + geom_density(aes(x=turnout, color=VoteCate))
ggplot(df) + geom_density(aes(x=turnout)) + geom_density(aes(x=turnout, color=VoteCate))

```

## Voter turnout and median household county income


```{r Graphs Turnout and Income}

ggplot(df) + geom_density(aes(x=turnout, color=VoteCate)) + 
    facet_grid(IncomeLevels~.) +
    labs(title="Restrictive Voting Laws Suppress Low-Income Turnout",
         subtitle="Distribution of 2016 General Election Turnout by County",
         caption="Counties are grouped by income, states are categorized based on the ease of voting relative to number of early voting days, mail-in ballot rules, and voter ID laws.") +
    theme_classic +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

```


```{r Graphs Drafts Income eval=FALSE, include=FALSE}

ggplot(df, aes(IncomeLevels, turnout, color=VoteByMail)) + geom_boxplot()

ggplot(df, aes(IncomeLevels, turnout, color=EarlyVoting)) + geom_boxplot()
ggplot(df, aes(IncomeLevels, turnout, color=factor(IDScore))) + geom_boxplot()

df %>% filter(IncomeLevels == "Low") %>% 
    ggplot() + geom_density(aes(x=turnout, color=VoteCate))

df %>% group_by(IncomeLevels, VoteByMail) %>%
    summarise(Turnout = sum(total_votes) / sum(X18.and.over.population)) %>%
    write_csv("Income_by_Mail_Means.csv")

summarise(df, Turnout = sum(total_votes) / sum(X18.and.over.population))



```

