---
title: "Voter Disenfranchisement: How laws stop eligible voters"
output:
  slidy_presentation: default
---


We will be looking at how citizens of voting age can be, and are being disenfranchised in the United States of America from these three points of view - the citizens and groups that are being adversely affected by disenfranchisement (Who), how location plays into it (the Where) and the temporal concerns that contribute to it (the When). 

```{r Libraries, include=FALSE}

library(tidyverse)

```

## Electoral and temporal concerns contributing towards disenfranchisement

Our main data source for this project comes from 2016 US Census.  We downloaded multiple files, each with various demographic information such as race, income, disability and employment. We merged these together along with results from the 2016 election to use as our main data file.

Information on early voting, abseentee ballots and other mail-in ballots came from the National Council of State Legislatures. These were html tables that we entered into .csv files to use in our analysis.

```{r Load Data, message=FALSE, warning=FALSE}

census <- read.csv("Exploratory_Analysis/census_with_election_data.csv") %>% 
    mutate(turnout = (total_votes/X18.and.over.population)*100)

laws <- read.csv("Voting Laws by State Revised.csv")

df <- left_join(census, laws, by=c('state.name'='State'))

df <- df[complete.cases(df),]

```

```{r New Vars message=FALSE, warning=FALSE}

df$EarlyVoting = case_when(df$EarlyVotingDays == 0 ~ "None",
                            df$EarlyVotingDays >= 18 ~ "18 to 45",
                            TRUE ~ "3 to 15")

df$bcollar <- df$Service + df$Natural.resources..construction..and.maintenance.occupations + df$Production..transportation..and.material.moving.occupations

df$wcollar <- df$Management..business..science..and.arts.occupations+ df$Sales.and.office.occupations

df$MailScore <- case_when(df$VoteByMail == "All Mail" ~ 1,
                         df$VoteByMail == "Any Absentee" ~ 2,
                         df$VoteByMail == "Excuse Only" ~ 3)

df$EarlyScore <- case_when(df$EarlyVoting == "None" ~ 3,
                          df$EarlyVoting == "3 to 15" ~ 2,
                          df$EarlyVoting == "18 to 45" ~ 1)

df$IDScore <- case_when(df$VoterID == "No ID" ~ 1,
                       df$VoterID == "Non Photo ID" ~ 2,
                       df$VoterID == "Strict Non-Photo" ~ 2,
                       df$VoterID == "Photo ID" ~ 3,
                       df$VoterID == "Strict Photo ID" ~ 3)

df$VoterScore <- df$MailScore + df$EarlyScore + df$IDScore

df$VoteCate <- case_when(df$VoterScore <= 4 ~ "Easy",
                         df$VoterScore >= 8 ~ "Hard",
                         TRUE ~ "Neutral")

df$DiverseCounty <- case_when(df$White <= .65 ~ "Yes",
                              TRUE ~ "No")

df$IncomeLevels <- case_when(df$Median.household.income < 40000 ~ "Low",
                             df$Median.household.income >= 55000 ~ "High",
                             TRUE ~ "Mid")

```


## County Distribution of Voter Turnout


```{r Graphs Overall Turnout Density}

ggplot(df, aes(x=turnout, group=VoteCate)) + 
    geom_density() +
    xlim(0,100) +
    labs(x="County voters as percentage of 18 years and older population",
         title="Flexible Voting Laws Increase Turnout",
         subtitle="In the 2016 general election, ease of voting is determined by restrictiveness of voter ID laws, number of early voting days, and vote by mail provisions.",
         caption="Sources: Population and election data from the U.S. Census Bureau. Voting laws from the National Council of State Legislatures. Excludes Alaska.") +
    annotate(geom = "text", x = c(46,74), y = c(.04,.045), label = c("Strict", "Flexible") , color="black", size=3) +
    theme_minimal() +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          legend.position="none",
          panel.grid.minor=element_blank(),
          panel.grid.major=element_blank())

#ggsave("CountyTurnout.png",width=10,height=6)

```


## Voter turnout in low income counties

```{r Graphs Turnout and Income}

df %>% filter(IncomeLevels == "Low") %>% ggplot() + 
    geom_density(aes(x=turnout, group=VoteCate)) + 
    xlim(0,100) +
    labs(x="County voters as percentage of 18 years and older population",
         title="Restrictive Voting Laws Suppress Low-Income Turnout",
         subtitle="In the 2016 general election, low income counties see lower turnout when early voting days and mail-in ballots are curtailed and photo IDs are required.",
         caption="Sources: Population and election data the from U.S. Census Bureau. Voting laws from the National Council of State Legislatures. Excludes Alaska.") +
    annotate(geom = "text", x = c(47,72), y = c(.051,.052), label = c("Strict", "Flexible") , color="black", size=3) +    theme_minimal() +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          legend.position="none",
          panel.grid.minor=element_blank(),
          panel.grid.major=element_blank())

#ggsave("LowIncomeTurnout.png",width=10,height=6)

```

## Voter turnout in high income counties

```{r Graphs Drafts Income eval=FALSE, include=FALSE}

df %>% filter(IncomeLevels == "High") %>% ggplot() + 
    geom_density(aes(x=turnout, group=VoteCate)) + 
    xlim(0,100) +
    labs(x="County voters as percentage of 18 years and older population",
         title="ADD CLAIM HERE",
         subtitle="ADD DETAIL HERE",
         caption="Sources: Population and election data the from U.S. Census Bureau. Voting laws from the National Council of State Legislatures. Excludes Alaska.") +
    annotate(geom = "text", x = c(47,72), y = c(.051,.052), label = c("Strict", "Flexible") , color="black", size=3) +    theme_minimal() +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          legend.position="none",
          panel.grid.minor=element_blank(),


```

## Voter turnout and Americans withe disabilities

From this graph we see the distribution of the turnout rates and the percentage of people with disability. Polling place inaccessibility may contribute to the lower voter turnout, both directly by making voting more difficult for people with disabilities, and indirectly by sending the message that people with disabilities are not expected to participate in the political sphere. We explore the role of polling place inaccessibility by examining voter turnout and reports of voting
difficulties among people with disabilities in US elections. We have tried to show it for all voters group, any voters and Excused only Group.

```{r Graphs Draft Turnout Disability}

#this needs to be redone to categorize these by disability

ggplot(df) + 
    geom_smooth(aes(Percent.With.a.disability,turnout, color=VoteCate)) +
    geom_smooth(aes(Percent.With.a.disability,turnout))
    


ggplot(df) + 
    geom_smooth(aes(Percent.With.a.disability,turnout, color=VoteByMail)) +
    geom_smooth(aes(Percent.With.a.disability,turnout))
    

ggplot(df) + 
    geom_smooth(aes(Percent.With.a.disability,turnout, color=EarlyVoting)) +
    geom_smooth(aes(Percent.With.a.disability,turnout))
    

```



